Causal Impact of Masks, Policies, Behavior on Early Covid-19 Pandemic in the U.S.

### 2020-09-09 tex/covidEffectOfNPI-revision.tex
tiny  update

Finally, our causal model is framed using the language of structural equations models and causal diagrams of econometrics (\cite{pwright, haavelmo:1944, tinbergen:1940, wold:1954, pearl:biometrika}) and genetics \citep{wright1923},\footnote{The father and son, P. Wright (economist) and S. Wright (geneticist) collaborated to develop structural equation models and causal path diagrams; P. Wright's key work represented supply-demand system as a directed acyclical graph and established its identification using exclusion restrictions on instrumental variables. We view our work as following this classical tradition.} with natural unfolding potential/structural outcomes representation \citep{rubin1974,tinbergen1930,neyman:PO,imbens_rubin_2015}. The work on causal graphs has been modernized and put on rigorous footing by  \cite{pearl:biometrika} and others (\cite{pearl:robins}, \cite{pearl:causality}, \cite{pearl:why}, \cite{white:chalak},  \cite{heckman:pinto}, \cite{peters2020book}, and \cite{hernanrobins2020book}), with major applications taking place in computer science, epidemiology, and econometrics. The particular causal diagram we use has several ``mediation" components, where variables affect outcomes directly and indirectly through other variables called mediators; these structures go back at least to \citet[][see Figure 6]{wright1923}; see, e.g., \cite{baron1986}, \cite{Hines2020}, \cite{richardson:mediation} for modern treatments.

@@

To estimate the coefficient of the $j$-th policy variable, $P_{it}^j$, using Double Machine Learning of \cite{chernozhukov18}, we consider a version of (\ref{eq:R4}) as follows:
\begin{align*}
Y_{i,t+\ell} & = P_{it}^j \theta_0 + \phi' X_{it} + g_0(W_{it}) + \xi_{it}, && E[\xi_{it}|P_{it}^j,X_{it},W_{it}]=0,\\
P_{it}^j &= \psi' X_{it} +m_0(W_{it}) +V_{it},&& E[V_{it}|X_{it},W_{it}]=0,
\end{align*}
where $X_{it}=((P_{it}^{-j})',B_{it}',I_{it}')'$ collects  policy variables except for the $j$-th policy variables $P_{it}^{-j}$, behavior variables $B_{it}$, and information variables $I_{it}$. The confounding factors $W_{it}$ affect the policy variable $P_{it}^j$ via the function $m_0(W)$ and the outcome variable $Y_{i,t+\ell}$ via the function $g_0(W)$. We apply Lasso  or  Random Forests to estimate $g_0(W)$  for dimension reductions or for capturing non-linearity while the coefficients of $X_{it}$ are estimated under linearity without imposing any dimension reductions.  We omit the details of estimation procedure here. Please see the discussion in Example 1.1 of  \cite{chernozhukov18} for reference.

>>>>

Finally, our causal model is framed using the language of structural equations models and causal diagrams of econometrics (\cite{pwright, haavelmo:1944, tinbergen:1940, wold:1954, pearl:biometrika}) and genetics \citep{wright1923},\footnote{The father and son, P. Wright (economist) and S. Wright (geneticist) collaborated to develop structural equation models and causal path diagrams; P. Wright's key work represented supply-demand system as a directed acyclical graph and established its identification using exclusion restrictions on instrumental variables. We view our work as following this classical tradition.} with natural unfolding potential/structural outcomes representation \citep{rubin1974,tinbergen1930,neyman:PO,imbens_rubin_2015}. The work on causal graphs has been modernized and put on rigorous footing by  \cite{pearl:biometrika} and others (\cite{pearl:robins,pearl:causality, pearl:why,white:chalak,heckman:pinto,hunermund,richardson:mediation,peters2020book,bareinboim:H,hernanrobins2020book}), with major applications taking place especially in computer science and epidemiology. The particular causal diagram we use has several ``mediation" components, where variables affect outcomes directly and indirectly through other variables called mediators; these structures go back at least to \citet[][see Figure 6]{wright1923}; see, e.g., \cite{baron1986}, \cite{Hines2020}, \cite{richardson:mediation} for modern treatments.

@@

To estimate the coefficient of the $j$-th policy variable, $P_{it}^j$, using double machine learning (\cite{chernozhukov18}), we consider a version of (\ref{eq:R4}) as follows:
\begin{align*}
Y_{i,t+\ell} & = P_{it}^j \theta_0 + \phi' X_{it} + g_0(W_{it}) + \xi_{it}, && E[\xi_{it}|P_{it}^j,X_{it},W_{it}]=0,\\
P_{it}^j &= \psi' X_{it} +m_0(W_{it}) +V_{it},&& E[V_{it}|X_{it},W_{it}]=0,
\end{align*}
where $X_{it}=((P_{it}^{-j})',B_{it}',I_{it}')'$ collects  policy variables except for the $j$-th policy variables $P_{it}^{-j}$, behavior variables $B_{it}$, and information variables $I_{it}$. The confounding factors $W_{it}$ affect the policy variable $P_{it}^j$ via the function $m_0(W)$ and the outcome variable $Y_{i,t+\ell}$ via the function $g_0(W)$. We apply Lasso  or  Random Forests to estimate $g_0(W)$  for dimension reductions or for capturing non-linearity while the coefficients of $X_{it}$ are estimated under linearity without imposing any dimension reductions.  We omit the details of estimation procedure here. Please see the discussion in Example 1.1 of  \cite{chernozhukov18} for reference.
